---
title: "Lessons Learned: Cars Dataset"
author: "Gerardo Parrello"
---

### Acquiring Data
#### Load data
```{r}
library("readr")
cars <- read.csv("./data/cars.csv")
```

#### Preprocess
- Change attribute names.
- Set data type to character for brand.
```{r}
names(cars) <- c("brand","speed","distance")
cars$brand <- as.character(cars$brand)
```

### Creating Train and Test Sets
#### Compute set sizes
```{r}
trainPct <- 0.8
trainSize <- round(nrow(cars)*trainPct)
testSize <- nrow(cars) - trainSize
```
We use `r trainPct*100`% of our data to train our model and the rest to test it. Our final set sizes are `r trainSize` and `r testSize` rows for training and testing, respectively.

#### Build sets
```{r}
seed <- 8888
set.seed(seed)
trainIndex <- sample(seq_len(nrow(cars)), size=trainSize)
```
Setting `r seed` as seed, we random sample the whole dataset for `r trainSize` indexes.

```{r}
trainSet <- cars[trainIndex,]
testSet <- cars[-trainIndex,]
```
We then filter the dataset with these indexes, building a new dataset with the indexes that match (trainSet) and another with the opposite, those that do not match (testSet).

### Modeling
#### Train Linear Model
```{r}
linear <- lm(distance~ speed, trainSet)
```

#### Apply Model to Test Set
```{r}
predictions <- predict(linear, testSet)
```

### Results
#### Plot Results
```{r}
plot(testSet$speed, predictions, col="Blue", xlab="Speed", ylab="Distance", main="Predicted and Real Distances")
points(testSet$speed, testSet$distance, col="Red")
abline(linear, col="Blue")
```

