---
title: 'Lessons Learned: Cars Dataset'
author: "Gerardo Parrello"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

### Load Libraries 
```{r}
library("readr")
library("rbokeh")
```

### Set Parameters
```{r}
plot_scale = .9
plot_width = 987 * plot_scale
plot_height = 610 * plot_scale
```

### Acquiring Data
#### Load Data
We have come to learn that the dataset contains the results of tests done in 1920, measuring the breaking distance and initial speed for several cars.
```{r}
cars <- read.csv("./data/cars.csv")
```

#### Find Info
```{r}
summary(cars)
str(cars)
```

#### Preprocess
- Change attribute names.
- Set data type to character for brand.
```{r}
names(cars) <- c("brand","speed","distance")
cars$brand <- as.character(cars$brand)
```

#### Show Data
```{r}
figure(width = plot_width,
       height = plot_height,
       title="Cars dataset") %>%
  ly_points(speed,
            distance,
            data = cars,
            hover = c(brand, speed, distance))
bins <- round(nrow(cars)/5)
figure(width = plot_width,
       height = plot_height,
       xlab="Speed",
       title="Speed histogram") %>%
  ly_hist(cars$speed, freq=FALSE, breaks=bins) %>%
  # don't care about absolute frequency, just the shape
  ly_density(cars$speed)
figure(width = plot_width,
       height = plot_height,
       lab="Distance",
       title="Distance histogram") %>%
  ly_hist(cars$distance, freq=FALSE, breaks=bins) %>%
  # don't care about absolute frequency, just the shape
  ly_density(cars$distance)
```

#### Remove Outliers
```{r}
cars_outliers <- cars[which(cars$distance > 100), ]
cars <- cars[which(cars$distance <= 100), ]
# the comma is important to select all columns from the dataframe
figure(width = plot_width,
       height = plot_height,
       title="Outliers") %>%
  ly_points(speed, distance,
            data = cars,
            hover = c(brand, speed, distance),
            legend="Original") %>%
  ly_points(speed,
            distance,
            data = cars_outliers,
            hover = c(brand, speed, distance),
            color="red",
            legend="Outlier")
```


### Creating Train and Test Sets
#### Compute set sizes
```{r}
trainPct <- 0.75
trainSize <- round(nrow(cars)*trainPct)
testSize <- nrow(cars) - trainSize
```
We use `r trainPct*100`% of our data to train our model and the rest to test it. Our final set sizes are `r trainSize` and `r testSize` rows for training and testing, respectively.

#### Build Sets
```{r}
seed <- 123
set.seed(seed)
trainIndex <- sample(seq_len(nrow(cars)), size=trainSize)
```
Setting `r seed` as seed, we random sample the whole dataset for `r trainSize` indexes.

```{r}
trainSet <- cars[trainIndex,]
testSet <- cars[-trainIndex,]
```
We then filter the dataset with these indexes, building a new dataset with the indexes that match (trainSet) and another with the opposite, those that do not match (testSet).

### Modeling
#### Train Linear Model
We create a model of the type $x(v)=m\cdot v+b$ where $x$ is distance and $v$ is speed.<!-- This will obviously fit our data since $v=\frac{\partial x}{\partial t}$-->
```{r}
linear <- lm(distance~ speed, trainSet)
summary(linear)
```

#### Apply Model to Train Set
```{r}
trainSet$pred_dist <- predict(linear, trainSet)
trainSet$error <- scale(
  trainSet$pred_dist - trainSet$distance
  ) # calculate normalized error
```

#### Plot Errors
```{r}
figure(width = plot_width,
       height = plot_height,
       title="Errors on train set prediction") %>%
  ly_points(speed,
            error,
            data = trainSet,
            hover = c(speed, error),
            color="green",
            glyph="triangle")
```

### Results
#### Apply Model to Test Set
```{r}
testSet$pred_dist <- predict(linear, testSet)
testSet$error <- scale(
  testSet$pred_dist - testSet$distance
  ) # calculate normalized error
```

#### Plot Results
```{r}
figure(width = plot_width,
       height = plot_height,
       title="Applied model on test set") %>%
  ly_points(speed,
            distance,
            data = testSet,
            hover = c(brand, speed, distance)) %>%
  ly_points(speed,
            pred_dist,
            data = testSet,
            hover = c(brand, speed, pred_dist),
            color="red") %>%
  ly_abline(linear, color="red", legend="Linear Model")
```

#### Plot Errors
```{r}
figure(width = plot_width,
       height = plot_height,
       title="Errors on test set prediction") %>%
  ly_points(speed,
            error,
            data = testSet,
            hover = c(speed, error),
            color="green", glyph="triangle")
```

