---
title: "Lessons Learned: Iris Dataset"
author: "Gerardo Parrello"
---
### Load Libraries 
```{r}
library("readr")
library("rbokeh")
```

### Set Parameters
```{r}
plot_scale = .9
plot_width = 987 * plot_scale
plot_height = 610 * plot_scale
```

### Acquiring Data
#### Load data
```{r}
IrisDataset <- read.csv("./data/iris.csv")
```

#### Find Info
```{r}
summary(IrisDataset)
str(IrisDataset)
```

#### Show Data
```{r}
figure(width = plot_width, height = plot_height, xlab="Petal Length", ylab="Petal Width", title="Iris dataset") %>%
  ly_points(Petal.Length, Petal.Width, data = IrisDataset, hover = c(Species, Petal.Length, Petal.Width), color=Species)
figure(width = plot_width, height = plot_height, xlab="Petal Width", title="Petal width histogram") %>%
  ly_hist(IrisDataset$Petal.Width, freq=FALSE) %>% # don't care about absolute frequency, just the shape
  ly_density(IrisDataset$Petal.Width)
figure(width = plot_width, height = plot_height, xlab="Petal Length", title="Petal height histogram") %>%
  ly_hist(IrisDataset$Petal.Length, freq=FALSE) %>% # don't care about absolute frequency, just the shape
  ly_density(IrisDataset$Petal.Length)
```

### Creating Train and Test Sets
#### Compute Set Sizes
```{r}
trainPct <- 0.8
datasetSize <- nrow(IrisDataset)
trainSize <- round(datasetSize * trainPct)
testSize <- datasetSize - trainSize
```
We use `r trainPct*100`% of our data to train our model and the rest to test it. Since our dataset is `r datasetSize` rows long, our final set sizes are `r trainSize` and `r testSize` rows for training and testing, respectively.

#### Build Sets
```{r}
seed <- 123
set.seed(seed)
training_indices <- sample(seq_len(datasetSize), size=trainSize)
```
Setting `r seed` as seed, we random sample the whole dataset for `r trainSize` indexes.

```{r}
trainSet <- IrisDataset[training_indices, ]
testSet <- IrisDataset[-training_indices, ]
```
We then filter the dataset with these indexes, building a new dataset with the indexes that match (trainSet) and another with the opposite, those that do not match (testSet).

### Modeling
#### Fit Linear Model
We create a model of the type $y=mx+b$.
```{r}
LinearModel <- lm(Petal.Width~ Petal.Length, trainSet)
summary(LinearModel)
```

#### Apply Model to Train Set
```{r}
trainSet$prediction <- predict(LinearModel, trainSet)
trainSet$error <- scale(trainSet$prediction - trainSet$Petal.Width)
```

#### Plot Errors
```{r}
figure(width = plot_width, height = plot_height, xlab="Petal Length", ylab="Error", title="Errors on train set prediction") %>%
  ly_points(Petal.Length, error, data = trainSet, hover = c(Petal.Length, error), color="purple", glyph="triangle")
```

### Results
#### Apply Model to Test Set
```{r}
testSet$prediction <- predict(LinearModel, testSet)
testSet$error <- scale(testSet$prediction - testSet$Petal.Width) # z-transform normalization of errors
```

#### Plot Results
```{r}
figure(width = plot_width, height = plot_height, xlab="Petal Length", ylab="Petal Width", title="Applied model on test set") %>%
  ly_points(Petal.Length, Petal.Width, data = testSet, hover = c(Species, Petal.Length, Petal.Width), color=Species) %>%
  ly_points(Petal.Length, prediction, data = testSet, hover = c(Petal.Length, prediction), color="red", glyph="square") %>%
  ly_abline(LinearModel, color="red", legend="Linear Model")
```

#### Plot Errors
```{r}
figure(width = plot_width, height = plot_height, xlab="Petal Length", ylab="Error", title="Errors on test set prediction") %>%
  ly_points(Petal.Length, error, data = testSet, hover = c(Petal.Length, error), color="purple", glyph="triangle")
```

