---
title: 'Lessons Learned: Iris Dataset'
author: "Gerardo Parrello"
output: pdf_document
---
### Load Libraries 
```{r}
library("readr")
library("rbokeh")
```

### Set Parameters
```{r}
plot_scale = .9
plot_width = 987 * plot_scale
plot_height = 610 * plot_scale
```

### Acquiring Data
#### Load data
```{r}
IrisDataset <- read.csv("./data/iris.csv")
```

#### Find Info
```{r}
summary(IrisDataset)
str(IrisDataset)
```

#### Show Data
```{r}
figure(width = plot_width,
       height = plot_height,
       xlab="Sepal Length",
       ylab="Sepal Width",
       title="Iris dataset") %>%
  ly_points(Sepal.Length,
            Sepal.Width,
            data = IrisDataset,
            hover = c(Species, Sepal.Length, Sepal.Width),
            color=Species)
```
There seems to be no pattern concerning sepal length and width.

But there is a relation for petal length and width.
```{r}
figure(width = plot_width,
       height = plot_height,
       xlab="Petal Length",
       ylab="Petal Width",
       title="Iris dataset") %>%
  ly_points(Petal.Length,
            Petal.Width,
            data = IrisDataset,
            hover = c(Species, Petal.Length, Petal.Width),
            color=Species)
figure(width = plot_width,
       height = plot_height,
       xlab="Petal Width",
       title="Petal width histogram") %>%
  ly_hist(IrisDataset$Petal.Width, freq=FALSE) %>%
  # don't care about absolute frequency, just the shape
  ly_density(IrisDataset$Petal.Width)
figure(width = plot_width,
       height = plot_height,
       xlab="Petal Length",
       title="Petal height histogram") %>%
  ly_hist(IrisDataset$Petal.Length, freq=FALSE) %>%
  # don't care about absolute frequency, just the shape
  ly_density(IrisDataset$Petal.Length)
```

The following code was adapted from [https://hafen.github.io/rbokeh](https://hafen.github.io/rbokeh).
```{r}
tools <- c("pan", "wheel_zoom", "box_zoom", "box_select", "reset")
attr_names <- names(IrisDataset)[2:5]
# Select attributes 2 to 5
nms <- expand.grid(attr_names, rev(attr_names), stringsAsFactors = FALSE)
# combinatoric function
splom_list <- vector("list", 16)
# 16 length vector because we have 4 attributes in matrix, 4x4 plots
for(ii in seq_len(nrow(nms))) {
  splom_list[[ii]] <- figure(width = plot_width/4,
                             height = plot_width/4,
                             tools = tools,
                             xlab = nms$Var1[ii],
                             ylab = nms$Var2[ii]) %>%
    ly_points(nms$Var1[ii],
              nms$Var2[ii],
              data = IrisDataset,
              color = Species,
              size = 5,
              legend = FALSE)
}
grid_plot(splom_list, ncol = 4,
          same_axes = TRUE,
          link_data = TRUE,
          height = plot_width,
          width = plot_width)
```


### Creating Train and Test Sets
#### Compute Set Sizes
```{r}
trainPct <- 0.8
datasetSize <- nrow(IrisDataset)
trainSize <- round(datasetSize * trainPct)
testSize <- datasetSize - trainSize
```
We use `r trainPct*100`% of our data to train our model and the rest to test it. Since our dataset is `r datasetSize` rows long, our final set sizes are `r trainSize` and `r testSize` rows for training and testing, respectively.

#### Build Sets
```{r}
seed <- 123
set.seed(seed)
training_indices <- sample(seq_len(datasetSize), size=trainSize)
```
Setting `r seed` as seed, we random sample the whole dataset for `r trainSize` indexes.

```{r}
trainSet <- IrisDataset[training_indices, ]
testSet <- IrisDataset[-training_indices, ]
```
We then filter the dataset with these indexes, building a new dataset with the indexes that match (trainSet) and another with the opposite, those that do not match (testSet).

### Modeling
#### Fit Linear Model
We create a model of the type $y=mx+b$.
```{r}
LinearModel <- lm(Petal.Width~ Petal.Length, trainSet)
summary(LinearModel)
```

#### Apply Model to Train Set
```{r}
trainSet$prediction <- predict(LinearModel, trainSet)
trainSet$error <- scale(trainSet$prediction - trainSet$Petal.Width)
```

#### Plot Errors
```{r}
figure(width = plot_width,
       height = plot_height,
       xlab="Petal Length",
       ylab="Error",
       title="Errors on train set prediction") %>%
  ly_points(Petal.Length,
            error,
            data = trainSet,
            hover = c(Petal.Length, error),
            color="purple",
            glyph="triangle")
```

### Results
#### Apply Model to Test Set
```{r}
testSet$prediction <- predict(LinearModel, testSet)
testSet$error <- scale(
  testSet$prediction - testSet$Petal.Width
  ) # z-transform normalization of errors
```

#### Plot Results
```{r}
figure(width = plot_width,
       height = plot_height,
       xlab="Petal Length",
       ylab="Petal Width",
       title="Applied model on test set") %>%
  ly_points(Petal.Length,
            Petal.Width,
            data = testSet,
            hover = c(Species, Petal.Length, Petal.Width),
            color=Species) %>%
  ly_points(Petal.Length,
            prediction,
            data = testSet,
            hover = c(Petal.Length, prediction),
            color="red",
            glyph="square") %>%
  ly_abline(LinearModel, color="red", legend="Linear Model")
```

#### Plot Errors
```{r}
figure(width = plot_width,
       height = plot_height,
       xlab="Petal Length",
       ylab="Error",
       title="Errors on test set prediction") %>%
  ly_points(Petal.Length, error,
            data = testSet,
            hover = c(Petal.Length, error),
            color="purple",
            glyph="triangle")
```

