---
title: "Lessons Learned: Iris Dataset"
author: "Gerardo Parrello"
---

### Acquiring Data
#### Load data
```{r}
library("readr")
IrisDataset <- read.csv("./data/iris.csv")
```

#### Find Info
```{r}
#attributes(IrisDataset)
#summary(IrisDataset)
#str(IrisDataset)
#names(IrisDataset)
#hist(IrisDataset$Species)
#plot(IrisDataset$Sepal.Length)
#qqnorm(IrisDataset)
```

#### Preprocess
 - Set data type to numeric for species attribute.
```{r}
IrisDataset$Species <- as.numeric(IrisDataset$Species)
```

### Creating Train and Test Sets
#### Compute set sizes
```{r}
trainPct <- 0.8
trainSize <- round(nrow(IrisDataset) * trainPct)
testSize <- nrow(IrisDataset) - trainSize
```
We use `r trainPct*100`% of our data to train our model and the rest to test it. Our final set sizes are `r trainSize` and `r testSize` rows for training and testing, respectively.


```{r}
seed <- 123
set.seed(seed)
training_indices <- sample(seq_len(nrow(IrisDataset)), size=trainSize)
```
Setting `r seed` as seed, we random sample the whole dataset for `r trainSize` indexes.

```{r}
trainSet <- IrisDataset[training_indices, ]
testSet <- IrisDataset[-training_indices, ]
```
We then filter the dataset with these indexes, building a new dataset with the indexes that match (trainSet) and another with the opposite, those that do not match (testSet).

#### Fit Linear Model
```{r}
LinearModel<- lm(Petal.Width~ Petal.Length, trainSet)
# summary(LinearModel)
```

#### Apply Model to Predict
```{r}
prediction <- predict(LinearModel, testSet)
prediction
```